---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r unzip, cache=TRUE}

x<-read.csv("activity.csv")
###load activity.csv file
 x$date<-as.Date(x$date,"%Y-%m-%d")
```


## What is mean total number of steps taken per day?

### 1.Calculate the total number of steps taken per day
```{r mean-day,}
tapply(x$steps, x$date, sum)->y
y
```
### 2.Make a histogram of the total number of steps taken each day
```{r histogram, }
hist(y, breaks = 20)
```
### 3.Calculate and report the mean and median of the total number of steps taken per day
```{r mean-median,}
median(y,na.rm=TRUE)
mean(y,na.rm=TRUE)
```



## What is the average daily activity pattern?

```{r averageday, }
na.omit(x)->x1
tapply(x1$steps,x1$interval,mean)->z
plot(unique(x1$interval),z,type="l")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r maxvalue,}
as.data.frame.table(z)->zdt
as.integer(as.character(zdt$Var1))->zdt$Var1
max(zdt$Freq)->max
zdt[zdt$Freq==max,1]
```


## Imputing missing values
### 1.Calculate and report the total number of missing values in the dataset

```{r na,}
na_count <-sapply(x, function(y) sum(length(which(is.na(y)))))
na_count
```

### 2. Devise a strategy for filling in all of the missing values in the dataset. 
```{r nafill,}
colnames(zdt)<-c("interval","steps")
na<-x[is.na(x$steps),]
newnadata<-merge(na, zdt, by=c("interval"))
newnadata<-newnadata[,-2]
colnames(newnadata)[3]<-"steps"

```
### 3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r newdata,}
rbind(x1,newnadata)->newdata
```
### 4.hist of mean and median of new and old na data
```{r newhist,}
tapply(newdata$steps,newdata$date, sum)->newsum
mean(newsum)
median(newsum)
hist(newsum)
hist(y,col = "black",add=T)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r newdatafinal,}
newdata$DayCategory <- ifelse(weekdays(newdata$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
```


```{r weekhist,}
library(plyr)
library(lattice)
final<- ddply(newdata, .(interval, DayCategory), summarize, Avg = mean(steps))
head(final)
##Plot data in a panel plot
xyplot(Avg~interval|DayCategory, final, type="l",  layout = c(1,2),
       main="Average Steps per Interval Based on Type of Day", 
       ylab="Average Number of Steps", xlab="Interval")

```




